---
title: "COVID-19 DataFest"
author: "Nagaprasad Rudrapatna and Jackson Muraika"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(stringr)
library(knitr)
library(tidyverse)
library(gridExtra)
library(leaps)
library(rms)
```

```{r load-data, message = FALSE}
whirlpool <- read_csv("data/whirlpool.csv")
lewiston_queenston <- read_csv("data/lewistonQueenston.csv")
rainbow <- read_csv("data/rainbow.csv")
bridges_total <- read_csv("data/combined_bridges1.csv")

bridges_total <- bridges_total %>%
  select(Month, Day, `Day of Week`, Auto, Bus, Truck, `RV/Limo`, `Total Vehicles`) %>%
  mutate(Month = as.factor(Month),
         Month = fct_relevel(Month, "January", "February", "March", "April"))

```

### Topic

We will investigate bridge traffic statistics from the Niagara Falls Bridge Commission and determine whether and how the COVID-19 pandemic has affected activity at the Lewiston-Queenston, Rainbow, and Whirlpool bridges. The Niagara Falls Bridges are the second-busiest port of entry between US and Canada (behind Kennedy International); thus, these traffic statistics offer a unique opportunity to examine how port crossing has been affected by the COVID-19 pandemic.

### Exploratory Data Analysis

We will begin by examining the aggregate data, which combines the traffic statistics from all three bridges -- Lewiston-Queenston, Rainbow, and Whirlpool.

```{r}
bridges_total1 <- bridges_total %>%
  group_by(Month) %>%
  summarise(total = sum(`Total Vehicles`)) %>%
  arrange(total)
bridges_total1
```

As we can see from the table above, bridge traffic was similar in January and February, with about 200,000 vehicles crossing in each month. However, when comparing February to March, we notice a large discrepancy in the number of vehicles -- roughly 80,000 less vehicles! Furthermore, comparing March to April reveals an even greater discrepancy -- almost 100,000 less vehicles! In sum, from January 2020 to April 2020, the number of vehicles crossing the Niagara Falls Bridges drastically decreased (approximately -180,000 vehicles).

Of course, this aligns with our expectations given the lockdown and social distancing guidelines which have been in place in March and April 2020.

```{r}
ggplot(data = bridges_total1, mapping = aes(x = Month, y = total)) +
  geom_point(color = "red") +
  labs(title = "Total Number of Crossing Vehicles",
       subtitle = "By Month",
       x = "Month",
       y = "Number of Crossing Vehicles")
```

Based on the scatterplot above, there seems to be a nonlinear relationship between month of the year and the number of vehicles crossing the bridges.

```{r}
bridges_total2 <- bridges_total %>%
  group_by(Month) %>%
  summarise(totalA = sum(Auto)) %>%
  arrange(totalA)
bridges_total2
```

```{r}
ggplot(data = bridges_total2, mapping = aes(x = Month, y = totalA)) +
  geom_point(color = "red") +
  labs(title = "Total Number of Crossing Autos",
       subtitle = "By Month",
       x = "Month",
       y = "Number of Crossing Autos")
```

```{r}
bridges_total3 <- bridges_total %>%
  group_by(Month) %>%
  summarise(totalB = sum(Bus)) %>%
  arrange(totalB)
bridges_total3
```

```{r}
ggplot(data = bridges_total3, mapping = aes(x = Month, y = totalB)) +
  geom_point(color = "red") +
  labs(title = "Total Number of Crossing Buses",
       subtitle = "By Month",
       x = "Month",
       y = "Number of Crossing Buses")
```

```{r}
bridges_total4 <- bridges_total %>%
  group_by(Month) %>%
  summarise(totalT = sum(Truck)) %>%
  arrange(totalT)
bridges_total4
```

```{r}
ggplot(data = bridges_total4, mapping = aes(x = Month, y = totalT)) +
  geom_point(color = "red") +
  labs(title = "Total Number of Crossing Trucks",
       subtitle = "By Month",
       x = "Month",
       y = "Number of Crossing Trucks")
```

```{r}
bridges_total5 <- bridges_total %>%
  group_by(Month) %>%
  summarise(totalRL = sum(`RV/Limo`)) %>%
  arrange(totalRL)
bridges_total5
```

```{r}
ggplot(data = bridges_total5, mapping = aes(x = Month, y = totalRL)) +
  geom_point(color = "red") +
  labs(title = "Total Number of Crossing RVs/Limos",
       subtitle = "By Month",
       x = "Month",
       y = "Number of Crossing RVs/Limos")
```

### Model

Now, we will attempt to fit a multiple linear regression model to explain the variation in the number of vehicles crossing the bridges based on the 

