---
title: "COVID-19 DataFest"
author: "Nagaprasad Rudrapatna and Jackson Muraika"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r set-up, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(stringr)
library(knitr)
```

```{r load-data, message = FALSE}
# whirlpool <- read_csv("data/whirlpool.csv")
# lewiston_queenston <- read_csv("data/lewistonQueenston.csv")
# rainbow <- read_csv("data/rainbow.csv")
bridges_total <- read_csv("data/combined_bridges1.csv")

bridges_total <- bridges_total %>%
  select(Month, Day, `Day of Week`, Auto, Bus, Truck, `RV/Limo`, `Total Vehicles`) %>%
  mutate(Month = as.factor(Month),
         Month = fct_relevel(Month, "January", "February", "March", "April"),
         `Day of Week` = as.factor(`Day of Week`),
         `Day of Week` = fct_relevel(`Day of Week`, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

```

### Topic

We will investigate bridge traffic statistics from the Niagara Falls Bridge Commission and determine whether and how the COVID-19 pandemic has affected activity at the Lewiston-Queenston, Rainbow, and Whirlpool bridges. The Niagara Falls Bridges are the second-busiest port of entry between US and Canada (behind Kennedy International); thus, these traffic statistics offer a unique opportunity to examine how port crossing has been affected by the COVID-19 pandemic.

### Exploratory Data Analysis

We will begin by examining the aggregate data, which combines the traffic statistics from all three bridges -- Lewiston-Queenston, Rainbow, and Whirlpool.

```{r}
bridges_total1 <- bridges_total %>%
  group_by(Month) %>%
  summarise(total = sum(`Total Vehicles`)) %>%
  arrange(total)
bridges_total1
```

As we can see from the table above, bridge traffic was similar in January and February, with about 200,000 vehicles crossing in each month. However, when comparing February to March, we notice a large discrepancy in the number of vehicles -- roughly 80,000 less vehicles! Furthermore, comparing March to April reveals an even greater discrepancy -- almost 100,000 less vehicles! In sum, from January 2020 to April 2020, the number of vehicles crossing the Niagara Falls Bridges drastically decreased (approximately -180,000 vehicles).

Of course, this aligns with our expectations given the lockdown and social distancing guidelines which have been in place in March and April 2020.

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = Auto)) +
  geom_violin(color = "orange") +
  labs(title = "Total Number of Daily Crossing Autos Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Number of Daily Crossing Autos")
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = Bus)) +
  geom_violin(color = "orange") +
  labs(title = "Total Number of Daily Crossing Buses Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Number of Daily Crossing Buses")
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = Truck)) +
  geom_violin(color = "orange") +
  labs(title = "Total Number of Daily Crossing Trucks Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Number of Daily Crossing Trucks")
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = `RV/Limo`)) +
  geom_violin(color = "orange") +
  labs(title = "Total Number of Daily Crossing RVs/Limos Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Number of Daily Crossing RVs/Limos")
```

```{r}
bridges_total2 <- bridges_total %>%
  group_by(Month) %>%
  summarise(totalA = sum(Auto)) %>%
  arrange(totalA)
bridges_total2
```

```{r prop}
bridges_total <- bridges_total %>%
  mutate(totalA = case_when(Month == "January" ~ 167326, Month == "February" ~ 170191, Month == "March" ~ 90570, Month == "April" ~ 4359))

bridges_total <- bridges_total %>%
  group_by(Month) %>%
  count(totalA) %>%
  mutate(avg_Auto = totalA / n)
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = avg_Auto)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Average Number of Daily Crossing Autos Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Average Number of Daily Crossing Autos")
```

```{r}
bridges_total <- bridges_total %>%
  mutate(totalB = case_when(Month == "January" ~ 338, Month == "February" ~ 364, Month == "March" ~ 197, Month == "April" ~ 2))

bridges_total <- bridges_total %>%
  group_by(Month) %>%
  count(totalB) %>%
  mutate(avg_Bus = totalB / nn)
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = avg_Bus)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Average Number of Daily Crossing Buses Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Average Number of Daily Crossing Buses")
```

```{r}
bridges_total <- bridges_total %>%
  mutate(totalT = case_when(Month == "January" ~ 30997, Month == "February" ~ 28931, Month == "March" ~ 30878, Month == "April" ~ 16020))

bridges_total <- bridges_total %>%
  group_by(Month) %>%
  count(totalT) %>%
  mutate(avg_Truck = totalT / n)
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = avg_Truck)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Average Number of Daily Crossing Trucks Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Average Number of Daily Crossing Trucks")
```

```{r}
bridges_total <- bridges_total %>%
  mutate(totalRV = case_when(Month == "January" ~ 126, Month == "February" ~ 108, Month == "March" ~ 116, Month == "April" ~ 28))

bridges_total <- bridges_total %>%
  group_by(Month) %>%
  count(totalRV) %>%
  mutate(avg_RV = totalRV / nn)
```

```{r}
ggplot(data = bridges_total, mapping = aes(x = Month, y = avg_RV)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Average Number of Daily Crossing RVs/Limos Per Month",
       subtitle = "On Niagara Falls Bridges",
       x = "Month",
       y = "Average Number of Daily Crossing RVs/Limos")
```

